Software Citation Principles

### 2016-07-23 sc-principles/software-citation-principles.tex
starting to try to improve intro to section 1
added some more text in response to R1 #9

@@

Note that some software which is or could be captured as part of data provenance may not be cited.
Citation is partly a record of software that is important to a research outcome\footnote{Citation can be used for many purposes, including for software: which software has been used in the work, which software has influenced the work, which software is the work superseding, which software is the work disproving, etc.}, where provenance is a record of all steps (including software) used to generated particular data within the research process.
This implies that for a data research product, provenance data will include some of the cited software.
Similarly, the software metadata that is recorded as part of data provenance should be a superset of the metadata recorded as part of software citation for the software that was used in the work.
The data recorded for reproducibility should also be a superset of the metadata recorded as part of software citation.
These statements may also be true for software products.
In general, we intend the software citation principles to cover the minimum of what is necessary for software citation for the purpose of software identification.
%Other use cases (e.g., provenance, reproducibility) may lead to additional requirements (i.e., enhanced metadata).
%\niemnote{Perhaps rephrase to ``
Some use cases related to citation (e.g., provenance, reproducibility) might have additional requirements beyond the basic metadata needed for citation, as Table~\ref{tab:use_cases} shows.
%''}
%\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}
%\dmnote{ Mention that the principles are covering what's necessary, not necessarily what is sufficient for a given use case. So perhaps we should discuss things like configurations or bugs? }

>>>>

\niemnote{How is this?}
The main contribution of this document are the software citation principles,
written fairly concisely in this section and discussed further later in the
document (\S\ref{sec:discussion}).

@@

Some software which is, or could be, captured as part of data provenance may not be cited.
Citation is partly a record of software important to a research outcome\footnote{Citation
can be used for many purposes, including for software: which software has been used in
the work, which software has influenced the work, which software is the work superseding,
which software is the work disproving, etc.}, where provenance is a record of all steps
(including software) used to generated particular data within the research process.
This implies that for a data research product, provenance data will include some of
the cited software.
Research results, including data, increasingly depend on software~\cite{Hannay:2009wp},
and thus may depend on the specific version
used~\cite{Sandve:2013gh,Wilson:2014aa}---furthermore, errors in software or
environment variations can affect results~\cite{Morin:2012hz,Soergel:2015aa}.
Similarly, the software metadata that is recorded as part of data provenance should be
a superset of the metadata recorded as part of software citation for the software that
was used in the work. The data recorded for reproducibility should also be a superset
of the metadata recorded as part of software citation.
These statements may also be true for software products.
In general, we intend the software citation principles to cover the minimum of what is
necessary for software citation for the purpose of software identification.
Some use cases related to citation (e.g., provenance, reproducibility) might have additional requirements beyond the basic metadata needed for citation, as Table~\ref{tab:use_cases} shows.
%\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}
%\dmnote{ Mention that the principles are covering what's necessary, not necessarily what is sufficient for a given use case. So perhaps we should discuss things like configurations or bugs? }
